---
title: "LING 416 Speech on the brain"
author: "Nan Wang"
date: "Spring 2026"

header-includes:
 - \usepackage{color}
 - \usepackage{framed}
 - \setlength{\fboxsep}{.8em}
 
output:
  pdf_document:
    toc: true
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  results = "markup", warning = FALSE, cache = TRUE,
  fig.align = "center", fig.width = 6.5)
```

```{r libraries, include = F}

```

# This document
This document hosts notes from LING416, Speech on the brain. 

Office hour: Friday Carlson 2nd floor 9-11am


# General takeaways

Todo: course requirements, grading, assignments, exams, etc.



# Lecture 3 (1/28)

Language family include Indo-European (English, German, French), Sino-Tibetan (Mandarin), Niger-Congo, Afro-Asiatic, etc.

How to decide whether two languages come from the same family? (1) sound correspondences (2) shared morphology (3) shared core vocabulary

Regarding language adaptation: 

Phonetics: 

- how to describe consonants: place, manner, voicing. Fricatives include /f/, /v/, /s/, /z/, /sh/, etc; stops include /p/, /t/, /k/; 

- how to describe vowels: height, backness, roundness





IPA charts: https://www.seeingspeech.ac.uk/ipa-charts/ 




Q{formallization of linguistics system through maths}




